### 2019-03-22 4_implementation.tex
read impl

One challenge in implementing Task Bench for such a wide variety of
systems is that the capabilities of the systems vary considerably. For
example, some systems are \emph{implicitly parallel}, and provide some
form of parallelism discovery from sequential programs, whereas others
are \emph{explicitly parallel} and require users to specify the
parallelism in the program. As a rule of thumb we aim for the Task
Bench implementations to be represenative of how actual applications
would be written in the respective systems. Places where the Task
Bench implementations may be unidiomatic are noted below.

@@

In all cases, members of the respective programming systems' teams
were consulted during the development and evaluation of the
corresponding Task Bench implementations. Their insight was invaluable
to ensuring that we provide the highest level of quality in the Task
Bench implementations for each system.

@@

PaRSEC~\cite{PARSEC13} is a task-based programming system which
supports two distinct programming models: \emph{parameterized task
  graphs} (PTG) and \emph{sequential task flow} (STF). In the PTG
model, the program consists of an analytical function which describes
the expansion of the task graph. PTG is therefore a more limited
programming model, but can potentially provide better performance
because the graph description is amenable to static analysis and can
be pruned efficiently for horizontally scalable execution.

In the STF model, tasks are enumerated in program order (by
dynamically executing the program), and dependencies between tasks are
identified automatically based on the arguments to tasks. A task
depends on another task if it reads data written by the other task,
and the data is copied automatically if the two tasks are executed on
different nodes. While the basic STF model is implicitly parallel
because the system is responsible for the discovery of parallelism, in
order to achieve horizontal scalability, the program is executed in a
SPMD fashion on all nodes, and the user is responsible for eliding
tasks that are not directly connected to those that are to be executed
on the current node.

Data in PaRSEC is stored in \emph{regions}, or dense arrays of
contiguous memory. Regions may be passed to tasks on other nodes, but
can only be written on the node where the data is originally
allocated. In our Task Bench implementation, each task writes to a
region as output, and these regions are passed as inputs to the
subsequent set of tasks. As in the OmpSs and OpenMP implementations,
the task launch code uses a switch statement as the PaRSEC API assumes
tasks have a fixed number of dependencies.

@@

Spark's programming model is based of \emph{functional parallelism}, a
generalization of the MapReduce programming
model~\cite{MapReduce04}. The core computational abstraction in Spark
are operators such as map, reduce, filter, join, etc. which by virtue
of being pure functions are trivially known to be
parallizable. Functions in Spark operate on \emph{resilient
  distributed datasets} (RDDs), which are globally-visible,
dynamically single-assignment data structures that cached in memory
and written to disk for durability. One of Spark's primary
optimizations therefore is the caching of RDDs to avoid unnecessary
disk traffic, resulting in substantial speedups over MapReduce.

@@

StarPU~\cite{StarPU11} is a task-based system that supports an STF
programming model similar to PaRSEC. Our Task Bench implementation in
StarPU is very similar to PaRSEC, including the use of a switch
statement to handle dynamic numbers of task dependencies, and manual
elision of tasks on unrelated nodes to provide horizontal scalability.

>>>>

One challenge in implementing Task Bench for such a wide variety of
systems is that the capabilities of the systems vary considerably. For
example, some systems are \emph{implicitly parallel}, and provide some
form of parallelism discovery from sequential programs, whereas others
are \emph{explicitly parallel} and require users to specify the
parallelism in the program. As a rule of thumb, we aim for the Task
Bench implementations to be representative of how actual applications
would be written in the respective systems. Places, where the Task
Bench implementations may be unidiomatic, are noted below.

@@

In all cases, members of the respective programming systems' teams
were consulted during the development and evaluation of the
corresponding Task Bench implementations. Their insight was invaluable
to ensure that we provide the highest level of quality in the Task
Bench implementations for each system.

@@

PaRSEC~\cite{PARSEC13} is a task-based programming system which
supports two distinct programming models: \emph{parameterized task
  graphs} (PTG) and \emph{dynamic task discovery} (DTD)~\cite{PARSEC_DTD}. 
PTG is a data flow model, where programmers write applications 
in a concise description of all data flows by describing tasks 
and how data flows between tasks. This creates a compressed algebraic
representation of the task graph, which is then transformed into C 
code by PaRSEC source-to-source compiler. PTG is therefore a more limited
programming model, but can potentially provide better performance
because the graph description is amenable to static analysis and can
be pruned efficiently for better horizontally scalable execution.

In the DTD model, tasks are enumerated in program order (by
dynamically executing the program), and dependencies between tasks are
identified automatically based on the input and output data of tasks. 
A task depends on another task if it reads data written by the other task,
and the data is copied automatically if the two tasks are executed on
different nodes. While the basic DTD model is implicitly parallel
because the system is responsible for the discovery of parallelism, in
order to achieve horizontal scalability, the program is executed in a
SPMD fashion on all nodes, and the user is responsible for eliding
tasks that are not directly connected to those that are to be executed
on the current node.

The PaRSEC build-in 2D block cyclic matrix is used to store data. 
A matrix is divided into M by N tiles where N equals to the width of
task graphs and M varies from 2 to the height of task graphs depends
on the optimization level of alleviating write-after-read depdenencies. 
Tiles are distributed across nodes in pre-defined patterns, 
and can be passed to tasks on other nodes, but
can only be written on the node where the data is originally
allocated. In our Task Bench implementation, each task writes to a
tile as output, and these tiles are passed as inputs to the
subsequent set of tasks. As in the OmpSs and OpenMP implementations,
the task launch code uses a switch statement as the PaRSEC API assumes
tasks have a fixed number of dependencies.

@@

Spark's programming model is based on \emph{functional parallelism}, a
generalization of the MapReduce programming
model~\cite{MapReduce04}. The core computational abstraction in Spark
are operators such as map, reduce, filter, join, etc. which by virtue
of being pure functions are trivially known to be
parallelizable. Functions in Spark operate on \emph{resilient
  distributed datasets} (RDDs), which are globally-visible,
dynamically single-assignment data structures that cached in memory
and written to disk for durability. One of Spark's primary
optimizations therefore is the caching of RDDs to avoid unnecessary
disk traffic, resulting in substantial speedups over MapReduce.

@@

StarPU~\cite{StarPU11} is a task-based system that supports an DTD
programming model similar to PaRSEC. Our Task Bench implementation in
StarPU is very similar to PaRSEC, including the use of a switch
statement to handle dynamic numbers of task dependencies, and manual
elision of tasks on unrelated nodes to provide horizontal scalability.
A simplified version of 2D block cyclic matrix from the Chameleon 
project~\cite{Chameleon} is used to describe data of tasks in 
our Task Bench implementation.
